\documentclass[a4paper,10pt, twocolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[usenames,dvipsnames]{color}
\usepackage{comment}
\usepackage{tikz}
\usepackage{verbatim}
\usetikzlibrary{arrows,shapes}
\usepackage{listings}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{tabularx}
\usepackage{enumitem}
\setlist{nolistsep}
\setlength{\parindent}{0cm}

%page boarders
\usepackage[top=3cm, bottom=3cm, left=2cm, right=2cm]{geometry}
\usepackage[style=numeric,backend=bibtex]{biblatex}
\addbibresource{../sources}
%\usepackage[style=mla,babel=hyphen,backend=biber]{biblatex}

\tikzstyle{vertex}=[circle,fill=black!25,minimum size=20pt,inner sep=0pt]
\tikzstyle{edge} = [draw,thick,->,>=latex,shorten >=1pt]
\tikzstyle{weight} = [font=\small]
\tikzstyle{selected edge} = [draw,line width=2pt,->,red!75,>=latex]
\tikzstyle{residual edge} = [draw,thick,->,blue!75,>=latex]
\tikzstyle{vertexE}=[circle,fill=black!25,minimum size=20pt,inner sep=0pt]

% Declare layers (for more convenience when drawing graphs)
\pgfdeclarelayer{background}
\pgfsetlayers{background,main}

\newtheorem{lemma}{Lemma}
\newtheorem{corollary}[lemma]{Corollary}
%\newtheorem{proof}[lemma]{Proof}

\title{Parallel Network Flow Algorithms \\ 
\large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Träff, lecture "Parallel Algorithms", 2012 winter term at TU Wien}}
\author{Martin Kalany, 0825673}

\begin{document}
\maketitle

\section{Abstract}
\label{sec:abstract}
We discuss the possibilities of parallelizing the problem of finding a maximum flow in a given flow network. Although intuitive sequential algorithms have been known for a long time, only the 1990s theoretical analysis of the problem was able to ascertain the inherently sequential nature of the problem, i.e., that no parallel algorithm that solves the problem in poly-logarithmic time with polynomially bounded resource consumption exists. Before this fundamental results, significant efforts were put into finding efficient parallel algorithms for this problem based on the PRAM model. We explain two fundamental results and compare them with later and improved developments.


\section{Introduction}
\label{sec:intro}
\emph{Maximum flow problem}\footnote{Henceforth called \lstinline|MAX-FLOW| for brevity}


\section{Network flows}
\label{sec:networkFlows}
A \textbf{flow network} \cite{ahuja93} is given by $N = (G,s,t,c)$, where
\begin{itemize}
	\item $G =(V,E)$ is a directed graph
    \item $s$ and $t$, $s \neq t$ are the source and terminal node
   	\item $c:E\rightarrow \mathbb{R}_0^{+}$ assigns a capacity $\forall a \in E$
   	\item $n=\lvert V\rvert$, $m=\lvert E\rvert$
\end{itemize}

Furthermore, the following assumptions are made:
\begin{itemize}
	\item $G$ is connected
	\item $G$ is simple, i.e., does not contain loops or parallel arcs
	\item $\nexists P(s,t)$ with infinite capacity
\end{itemize}

\medskip
$f:E \rightarrow \mathbb{R}_0^{+}$ is a \textbf{flow} if it satisfies:
\begin{itemize}
	\item \emph{Capacity constraints:} $f(e) \leq c(e)$ $\forall e \in E$
	\item \emph{Flow conservation:} 
	$ \sum\limits_{v \in V} f(u,v) =  0 \Leftrightarrow IN(f,v) = OUT(f,v)$ $\forall v \in V \setminus \{s,t\}$
	\item \emph{Skew symmetry:} $f(v,w) = -f(w,v)$
	\item \emph{Value of a flow:} $\lvert f\rvert = f(V,t)$ 
\end{itemize}

A flow $f$
\begin{itemize}
	\item is a \emph{maximum flow} if $\lvert f\rvert \geq \lvert f'\rvert$, for any other flow $f'$
	\item \emph{saturates} an arc e if $f(e) = c(e)$
	\item is a \emph{maximal (or blocking) flow} if every directed path P(s,t) contains at least one saturated arc
\end{itemize}
\medskip
The \emph{residual capacity} of $e \in V \times V$ w.r.t. a flow $f$ is defined as $c_r(e) = c(e) - f(e)$. $G_r = (V, E_r)$ is the \emph{residual network}, where $E_r = \left\{e \in V \times V \lvert c_r(e) > 0\right\}$. The set of \emph{admissible arcs} of a vertex $v$ is the set of all edges $e$ emanating from $v$ s.t. $c_r(e) > 0$.  A path $P$ from $s$ to $t$ in $G_r$ is called an \emph{augmenting path} if it consists of admissible arcs only and can be used to increase the flow $f$.

\subsection{Algorithm of Ford-Fulkerson}
\label{sec:fordfulkerson}
The algorithm of Ford and Fulkerson as shown in Algorithm \ref{algo:fordFulkerson} is an intuitive sequential algorithm that solves \lstinline|MAX-FLOW|: starting from a null flow (or any other valid flow), increase the flow along an augmenting path as long as such a path exists. This is done by finding a path from $s$ to $t$ in the residual network along which the flow can be increased. 

\begin{algorithm}
\caption{Ford-Fulkerson}
\label{algo:fordFulkerson}
\begin{algorithmic}[1]
	\Function{Ford-Fulkerson}{$N=(G=(V,E),s,t,c)$}
	\State $f(u,v) = 0$ $\forall e=(u,v) \in E$ \Comment{Start with a valid flow}
	\State  $G_r = G$ 
	\While{$\exists P(s,t) \in G_r: \forall (u,v) \in P(s,t): c_r(u, v) > 0$}
		\State $c_r(P) = min \{c_r(u,v): (u,v) \in P(s,t) \}$
		\State $f(u,v) = f(u,v) + c_r(P)$ \Comment{Increase the flow by as much flow as can be pushed along $P(s,t)$}
		\State $f(v,u) = f(v,u) + c_r(P)$ \Comment{Update the residual network}
	\EndWhile
	\EndFunction
\end{algorithmic}
\end{algorithm}

Figure \ref{fig:fordfulkerson1} depicts the flow network for which \lstinline|MAX-FLOW| is to be solved and the first augmenting path\footnote{The residual network is equal to the given network in the first step.}. Note that this path may be chosen arbitrarily. Figure \ref{fig:fordfulkerson2} shows the residual network resulting from the situation in Figure \ref{fig:fordfulkerson1} and a possible next augmenting path. 

Note that the algorithm of Ford and Fulkerson is restricted to integer capacities:  $c:E\rightarrow \mathbb{N}_0^{+}$ $\forall a \in E$. For a more detailed explanation and theoretical analysis, refer to \cite{ahuja93}.

\begin{figure}
\begin{center}
\begin{tikzpicture}[scale=1.2, auto,swap]
    % draw the vertices
	\foreach \pos/\name in {{(0,2)/s}, {(1,3)/v_1}, {(2.5,3)/v_2},
   	                        {(1,1)/v_3}, {(2.5,1)/v_4}, {(3.5,2)/t}}
    \node[vertex] (\name) at \pos {$\name$};
   	% Connect vertices with edges and draw weights
   	\foreach \source/ \dest /\weight in {s/v_1/2, s/v_3/8,v_3/v_1/5,
                                         v_1/v_2/7, v_4/v_2/15, v_3/v_4/5,
                                         v_2/t/2, v_4/t/5}
       	\path[edge] (\source) -- node[weight] {$\weight$} (\dest);
        	
       	%v_1/v_4/9
       	\path[edge] ([xshift= -2pt, yshift= 5pt] v_4.center) -- node[weight] {6}  ([xshift= 5pt, yshift= -2pt] v_1.center);
       	\path[edge] ([xshift = 2pt, yshift= -5pt] v_1.center) -- node[weight] {9}  ([xshift= -5pt, yshift= 2pt] v_4.center);
  
	% For convenience we use a background layer to highlight edges
   	% This way we don't have to worry about the highlighting covering
	% weight labels. 
	\begin{pgfonlayer}{background}
   	    \foreach \source /\dest  /\weight in {s/v_3/5,v_4/t/5}
	        \path[selected edge] (\source) -- node[weight,above] {$\weight$}(\dest);		
            \path[selected edge] ([xshift = 2pt, yshift= -5pt] v_1.center) -- node[weight,left] {5}  ([xshift= -5pt, yshift= 2pt] v_4.center);
            \path[selected edge] (v_3) -- node[weight,left] {$5$}(v_1);
   	\end{pgfonlayer}
\end{tikzpicture}
\end{center}
\caption{Algorithm of Ford-Fulkerson: Initial network (black) and first augmenting path (red)}
\label{fig:fordfulkerson1}
\end{figure}

\begin{figure}
\begin{center}
\begin{tikzpicture}[scale=1.2, auto,swap]
   	% draw the vertices
    \foreach \pos/\name in {{(0,2)/s}, {(1,3)/v_1}, {(2.5,3)/v_2},
    	                        {(1,1)/v_3}, {(2.5,1)/v_4}, {(3.5,2)/t}}
    \node[vertex] (\name) at \pos {$\name$};
   	% Connect vertices with edges and draw weights
   	\foreach \source/ \dest /\weight in {s/v_1/2, v_1/v_3/5, t/v_4/5,
                                         v_1/v_2/7, v_4/v_2/15, v_3/v_4/5,
                                         v_2/t/2}
       	\path[residual edge] (\source) -- node[weight] {$\weight$} (\dest);

        	%s/v_3/8,
       	\path[residual edge] ([xshift= -2pt, yshift= 5pt] v_3.center) -- node[weight] {5}  ([xshift= 5pt, yshift= -2pt] s.center);
       	\path[residual edge] ([xshift = 2pt, yshift= -5pt] s.center) -- node[weight] {3}  ([xshift= -5pt, yshift= 2pt] v_3.center);
        	%v_1/v_4/9
       	\path[residual edge] ([xshift= -2pt, yshift= 5pt] v_4.center) -- node[weight] {11}  ([xshift= 5pt, yshift= -2pt] v_1.center);
       	\path[residual edge] ([xshift = 2pt, yshift= -5pt] v_1.center) -- node[weight] {4}  ([xshift= -5pt, yshift= 2pt] v_4.center);       	
        	
       	\begin{pgfonlayer}{background}
   	    \foreach \source /\dest  /\weight in {s/v_1/2,v_1/v_2/2,v_2/t/2}
            \path[selected edge] (\source) -- node[weight,above] {$\weight$}(\dest);			            
   	\end{pgfonlayer}
\end{tikzpicture}
\end{center}
\caption{Algorithm of Ford-Fulkerson: Residual network (blue) and augmenting path (red)}
\label{fig:fordfulkerson2}
\end{figure}	

\section{Model of Computation}
\label{sec:model}
We use the \textbf{PRAM} Model\footnote{Parallel Random Access Machine, as defined in the lecture or in \cite{papa95}} as the model of computation. Specifically, a \textbf{MIMD} - Multiple Instructions Multiple Data - variant is used, where each processor $p$ executes a distinct set of instructions on distinct data. Instructions are executed in lock-step and any necessary padding of instructions is assumed to be done automatically. 

\section{Computitional Complexity}
\label{sec:cc}
The algorithmic complexity of the \emph{Ford-Fulkerson} algorithm is given as $O((n+m)*f_{max})$\footnote{$f_{max}$ is from here on used to denote the maximum flow} \cite{ahuja93,papa95}. This was improved by the \emph{Edmonds-Karp} algorithm to $O(n*m^2)$, by using shortest augmenting paths instead of arbitrarily chosen ones. Clearly, the \emph{maximum flow problem} can be solved in polynomial time with a sequential algorithm and is thus $\in P$.

Informally, algorithms may be \emph{efficiently parallelizeable} or \emph{inherently sequential}. The complexity class \textbf{NC} \cite{papa95} ("Nick's Class") contains all problems solvable in $O(log^{k_1}(n))$ time and $O(n^{k_2})$ total work. We give a useful alternate definition:  A language that is decided by PRAM in $O(log^{k_1}(n))$ time steps with $O(n^{k_2})$ processors available at each step is in $NC$. Problems $\in NC$ thus are solvable  in poly-logarithmic time requiring polynomial work (or resources). Problems $P \setminus NC$ are dubbed "inherently sequential", i.e., not parallelizeable efficiently. 

Clearly, $NC \subseteq P$, but whether $NC \subset P$ or $NC = P$ is still unknown. It follows that we do not know whether or not inherently sequential problems actually exist. The most likely problems to be hard to parallelize are P-complete problems. Since any problem $A \in P$ can be reduced to a P-complete problem $B$, a parallel algorithm that solves $B$ within poly-logarithmic time bounds would immediately imply that for all problems $\in P$ a parallel algorithm $\in NC$ exists \cite{papa95}.

As an example, the Ford-Fulkerson algorithm can not be parallelized trivially. Although constructing the residual network can be done in constant time $O(1)$ by assigning a processor to each edge $a \in E$ and augmenting paths can be found in $O(log^{2}n)$ time and $O(n^{2})$ work by a basic BFS search, the number of stages can not be reduced to less than $O(\sqrt(n))$ \cite{ahuja93}. Thus the total complexity, which is dominated by the number of stages, places the algorithm in $P \setminus NC$.
	
As shown in \cite{papa95}, \lstinline|MAX-FLOW| is P-complete and no parallel algorithm $\in NC$ is currently known for solving this problem. It is assumed that the problem is inherently sequential\footnote{assuming that $P \neq NP$.}.

\section{Dinitz' scheme}
\label{sec:dinitz}
$N = (V,E,s,t,c)$ is a \textbf{layered network} \cite{yossi81} if each vertex $v \in V$ has a layer number $l(v)$ s.t.
\begin{itemize}
	\item $l(s) = 0$ and $0 \leq l(v) \leq l(t)$ $\forall v \in V$
	\item $(e = u \rightarrow v) \in E  \Rightarrow l(v) - l(u) = 1$
\end{itemize}
Figure \ref{fig:dinitz} shows a layered network and it's underlying directed acyclic network.

\begin{figure}
\begin{center}
\begin{tikzpicture}[scale=1.2, auto,swap]
   	% draw the vertices
	\foreach \pos/\name/\layer in {{(0,2)/s/0}, {(1,3)/v_1/1}, {(2.5,3)/v_2/2},
    	                        {(1,1)/v_3/1}, {(2.5,1)/v_4/2}, {(3.5,2)/t/3}}
    \node[vertex,label={[color=blue]80:$\layer$}] (\name) at \pos {$\name$};
        
    %\node[vertex,label={[color=blue]80:$+12$}] (v_1) at (1.5,3) {$v_1$};
    % Connect vertices with edges and draw weights
    \foreach \source/ \dest in {s/v_1, s/v_3,v_3/v_1,
                                        v_1/v_2, v_4/v_2, v_3/v_4,
                                        v_2/t, v_4/t}
     \path[edge] (\source) -- (\dest);
        	
     %v_1/v_4/9
     \path[edge] ([xshift= -2pt, yshift= 5pt] v_4.center) -- ([xshift= 5pt, yshift= -2pt] v_1.center);
     \path[edge] ([xshift = 2pt, yshift= -5pt] v_1.center) -- ([xshift= -5pt, yshift= 2pt] v_4.center);
  
	% For convenience we use a background layer to highlight edges
    % This way we don't have to worry about the highlighting covering
	% weight labels. 
	\begin{pgfonlayer}{background}
    	\foreach \source /\dest in {s/v_3,v_2/t,s/v_1,v_1/v_2,v_3/v_4,v_4/t}
	      	\path[selected edge, color=blue] (\source) -- (\dest);		
	    \path[selected edge, color=blue] ([xshift = 2pt, yshift= -5pt] v_1.center) -- ([xshift= -5pt, yshift= 2pt] v_4.center);
	\end{pgfonlayer}
\end{tikzpicture}
\end{center}
\caption{A layered network (in blue) with layer numbers and it's underlying network}
\label{fig:dinitz}
\end{figure}

\begin{corollary}
\textbf{Dinitz' scheme} \cite{dinitz70} \\
A maximum flow problem in a general network can be transformed into $O(n)$ maximal flow problems in layered networks.
\end{corollary}

A layered network can easily be constructed from a directed acyclic network by performing breath-first search in  $O(log^{2}n)$ time. It follows that an algorithm that finds a maximal/blocking flow in a layered network can solve \lstinline|MAX-FLOW| in $O(n)$ iterations. The basic scheme is given in Algorithm \ref{algo:dinitz}

\begin{algorithm}
\caption{Dinitz' scheme}
\label{algo:dinitz}
\begin{algorithmic}[1]
	\Function{MAX-FLOW}{$N=(V,E,s,t,c)$}
	\State start with some valid flow 
	\While{$l(t) \neq \infty$} %\Comment{\textcolor{OliveGreen}{$O(n)$}}	
		\State Construct residual network $G_r$ %\Comment{\textcolor{OliveGreen}{$O(n)$, $p=O(n)$}}
		\State Construct layered network $G_l$ from $G_r$ %\Comment{\textcolor{OliveGreen}}{$O(m/p + n)$}
		\State $f_l$ = MAX-FLOW($G_l$)  %\Comment{tbd}
		\State $f$ = $f$ + $f_l$ %\Comment{\textcolor{OliveGreen}{$O(n)$}}
	\EndWhile
	\EndFunction
\end{algorithmic}
\end{algorithm}

\section{Algorithm of Shiloach and Vishkin}
\label{sec:shiloach}
The Shiloach-Vishkin algorithm is based on Dinitz' scheme as shown in Section \ref{sec:dinitz}. This section focuses on the implementation of an algorithm to solve \lstinline|MAX-FLOW| for a given layered network.

The following notation will be used:
\begin{itemize}
	\item \lstinline|EXCESS(v)|: The amount of excessive flow at node $v$. \lstinline|EXCESS(v)| = $IN(f,v) - OUT(f,v) \geq 0$.
	\item \lstinline|UNBALANCED|: The set of vertices that hold some excess flow. \lstinline|UNBALANCED| = $\{v \in V \lvert EXCESS(v) > 0 \}$. The implementation uses a queue to maintain FIFO order on the set of unbalanced vertices.
	\item \lstinline|blocked|: A vertex v is blocked if for all it's outgoing edges $e_i = (v, u_i)$, $e_i$ is either saturated or $u_i$ is blocked\footnote{The sink vertex $t$ is never blocked}. 
\end{itemize}

The algorithm operates in \emph{pulses}: In the first pulse, all outgoing edges of $s$ or saturated: $\forall e_i = (s, v_i):$ $f(e_i) = c(e_i)$. The set of unbalanced vertices at the end of pulse 1 is thus the set of all neighbours of $s$: \lstinline|UNBALANCED| = $\{v_i\lvert \exists (s, v_i) \in E \}$. In each subsequent pulse, the unbalanced vertices try to push forward\footnote{i.e., to the next layer of the network} as much of their excessive flow as possible. Any remaining excessive flow will be returned backward. All balanced vertices remain idle during the pulses. The algorithm terminates as soon as no unbalanced vertices are available, implying that all flow reach the source $s$ or sink $t$.

Algorithm \ref{algo:sv} shows an abstract routine for solving \lstinline|MAX-FLOW| in a given layered network. In each pulse $i>1$, the \lstinline|while| loop in lines 4-13 is executed in parallel for each node $v \in UNBALANCED$. Parallel implementations of functions \lstinline|PUSH| and \lstinline|RETURN| are depicted in Algorithm \ref{algo:sv_push} and Algorithm \ref{algo:sv_return}, respectively and will be explained in detail in the remainder of this section. The algorithm maintains the number of the current pulse in $i$.

\begin{algorithm}
\caption{Shiloach-Vishkin}
\label{algo:sv}
\begin{algorithmic}[1]
	\Function{MAX-FLOW}{$G_l$}
		\State EXCESS(s) = $\Sigma_{v \in L_1}$ c(s$\rightarrow$v)
		\State PUSH(s, EXCESS(s))
		i = 1;
		\While{$\exists$ v $\in UNBALANCED$}
			\State i++
			\ForAll {$v \in UNBALANCED$}			
				\If{$v$ is not blocked}
					\State PUSH(v, EXCESS(v))
				\EndIf
			\EndFor
			\State RETURN(v, EXCESS(v))
		\EndWhile
	\EndFunction
\end{algorithmic}
\end{algorithm}

\subsection{Partial Sums trees}
\label{sec:sv_pstrees}
A \emph{partial sums tree}\footnote{Abbreviated PS-tree from here on.} is a complete binary tree with the leftmost $k$ leaves associated with $k$ numbers $a_1,...,a_k$. A partial sums tree is of height $h = \lceil log_2(k) \rceil$ and has $w = 2^{\lceil log_2(k) \rceil}$ leaves. The remaining $w - k$ rightmost leaves are associated with 0. Each internal node of a partial sums tree stores the sum of it's child nodes. Thus, the root node stores the sum of all leaves of the tree. Figure \ref{fig:pstree} provides an example.

\begin{figure}[H]
\begin{center}
  \begin{tikzpicture}[scale=0.8,label distance=-3mm,>= latex, node distance = 2cm,level/.style={sibling distance=40mm/#1}]
    \tikzstyle{every node}=[draw=none,fill=none, minimum size=5mm,font=\footnotesize];

   \node (pr) {$23$}
   	child {node (p13) {$17$}
	  	child {node (p9) {$8$} 
    	 	child {node (p1) {$5$}}
      		child {node (p2) {$3$}}
    		}
  		child {node (p10) {$9$} 
    	  	child {node (p3) {$2$}}
      		child {node (p4) {$7$}}
   		 }
   	}
   	child {node (p14) {$6$} 
   		child {node (p11) {$6$} 
      		child {node (p5) {$6$}}
		    child {node (p6) {$0$}}
    		}
   		child {node (p12) {$0$} 
      		child {node (p7) {$0$}}
      		child {node (p8) {$0$}}
      	}
    };
\end{tikzpicture}
\end{center}
\caption{A partial sums tree with $k=5$}
\label{fig:pstree}
\end{figure}

A \emph{flow quantum} is a 3-tuple $(p, e, f)$, where $p$ is the number of the pulse during which some flow of value $f$ reached a vertex $v$ via edge $e$.

For the parallel implementation of \lstinline|PUSH| and \lstinline|RETURN|, several PS-trees will be used. Four PS-trees are attached to every vertex $v \in V$:
\begin{itemize}
	\item T-OUT(v): Each active leave $a_i$ is associated with an outgoing edge $e_i = (v, w_i)$. $a_i = c(e_i) - f(e_i)$ ,i.e., $a_i$ stores the residual capacity of $e_i$. T-OUT(v) has $d_{out}(v)$\footnote{As usual, $d_{in}(v)$ and $d_{out}(v)$ designate the in- and out-degree of a vertex} active leaves.
	\item T-IN(v): Records flow quanta reaching $v$. T-IN(v) has $2n*d_{in}(v)$ active leaves to be able to record the pulse\footnote{As shown in Section \ref{sec:sv_analysis}, the number of pulses is bounded by $O(2n)$} and edge for each flow quantum that reaches $v$. 
	\item T-ACCESS(v): Each of the $d_{in}(v)$ active leaves is associated with one edge entering $v$. Used to coordinate access to $v$.
	\item T-SUM(v): Sums up the flow returned to $v$ in a given pulse. $d_{out}(v)$ active leaves. 
\end{itemize}
One PS-tree is attached to every arc $e \in E$:
\begin{itemize}
	\item T-EDGE(v): $2n$ active leaves, each of which is associated with one pulse to record the amount of flow returned on $e$ in a given pulse.
\end{itemize}

Let T be a PS-tree. Every node of T is represented by $T[height, i]$, where $height(root) = \lceil log_2(k) \rceil$ and $heigt(a_i) = 1$ , $\forall$ leaves $a_i$. Additionally, $T[root]$ designates the root node of T. We define the following operations on T:
\begin{itemize}
	\item CLEAR(i): The value of all nodes on the path $P(T[1,i],root)$ is set to 0. Note that executing this operation for less than all leaves will result in an invalid PS-tree. When executing the operation for all leaves, several processes will attempt to write to the same node. Thus a CRCW-PRAM is required.
	\item UPDATE(i,$a_i$): The value of the $i^{th}$ leave is set to $a_i$ and the values of all nodes on the path $P(T[1,i],root)$ are updated.
	\item SUM(i, $S_i$): Calculate the sum of leave nodes up to $a_i$. $S_i = a_1,...,a_i$.
	\item FIND($\alpha$, k, $\rho$): Given $\alpha$, find $k$ and $\rho$ s.t. $a_1+...+a_{k-1} < \alpha \leq a_1+...+a_k$ and $\rho = \alpha - (a_1+...+a_{k-1}$.
\end{itemize}  

\subsection{Parallel Implementation}
\label{sec:sv_parImpl}
We use the following processor assignment:
\begin{itemize}
	\item Each vertex $v \in V$ has a processor $P(v)$ assigned
	\item Each edge $e \in E$ has a processor $P(e)$ assigned
	\item Each flow quantum $Q$ has a processor $P(Q)$ assigned. In other words, each leaf in each of the $2n * d_{in}(v)$ T-IN(v) PS-trees has a processor $P(Q)$ assigned.
\end{itemize}

The total number of required processors is $O(nm)$, which is dominated by the assignment of one processor for each flow quantum. This large number of required processors can be reduced to $n$, which we will prove in Section \ref{sec:sv_nr_procs}. We will however use the above assignment of processors to present the essentials of the algorithm in cleaner way.   

In addition to the four PS-trees attached to each vertex $v$, $v$ keeps two pointers:
\begin{itemize}
	\item head(v): Points to the rightmost significant leave in T-in($v$). The rightmost significant (i.e., the rightmost leave with a value $\neq 0$) is associated with the flow quanta that was the last to reach $v$. head(v) will be used to return excess flow. 
	\item $k'$: The outgoing edges of $v$ are assumed to be ordered arbitrarily. $k'$ points to the first outgoing edge\footnote{the edge with the smallest index w.r.t. the given ordering} that is not yet saturated. In other words, $k'$ points to the next edge that can be utilized to push excess flow from $v$.  
\end{itemize}

In Algorithms \ref{algo:sv_init}, \ref{algo:sv_push} and \ref{algo:sv_return}, the notation \textcolor{blue}{P():} is used to indicate that the following instructions are carried out only by the specified processors. Algorithm \ref{algo:sv_init} shows the initialization routine executed at the beginning of each pulse. Each processor assigned to an edge $e_j = (v, w)$ updates the PS-tree T-OUT attached to $v$ s.t. it's leave $a_j$ is assigned the residual capacity of $e_j$. $f(e_j$) is set to 0. In the next step, each processor assigned to a vertex $v$ initializes it's pointers $head(v)$ and $k'(v)$.

\begin{algorithm}
\caption{Shiloach-Vishkin: INITIALIZE}
\label{algo:sv_init}
At beginning of each pulse:	
\begin{algorithmic}[1]
	\Function{INTIALIZE}{v}
		\State \textcolor{blue}{P($e_j = v \rightarrow w$):}
		\State UPDATE($j$, $c_r(e_j)$) T-OUT(v)
		\State f($e_j$) = 0
		\State \textcolor{blue}{P(v):}
		\State head(v) = 0	
		\State k'(v) = 1	
 	\EndFunction
\end{algorithmic}
\end{algorithm}

The \lstinline|PUSH| function is given in Algorithm \ref{algo:sv_push}. In lines 3-4, the amount of flow to be pushed in this pulse from node $v$ is calculated and the amount of excess flow remaining at node $v$ is updated. In line 5, the algorithm computes the number $k-1$ of outgoing edges that will be saturated in this pulse, starting with the edge pointed to by $k'$, as well as the amount of flow $\rho$ that will be pushed on the $k^{th}$ outgoing edge, without saturating that edge.

In the next step, all processors assigned to an edge $e_j$ s.t. $k'\leq j \leq$ will update their edge: In line 10, the flow of the $k^{th}$ edge is updated by an amount $\rho$; in line 12, $k-1$ edges are updated with as much flow as can be pushed through them. Lines 8-9 are used to count and order the processes that require access to node $w$, in order to store the flow quanta pushed to $w$ (line 16). In line 18, the \lstinline|head| pointer of node $w$ is advanced to the last recorded flow quantum. Lines 15 and 20 calculate the amount of excess flow at node $w$ (which is the amount of flow that was pushed to $w$ in this pulse). Finally, in lines 23-26 all vertices $v$ that have some remaining excess flow are blocked. Since no more flow can be pushed along their outgoing edges, the remaining excess flow will have to be returned.

\begin{algorithm}
\caption{Shiloach-Vishkin: PUSH}
\label{algo:sv_push}
\begin{algorithmic}[1]
	\Function{PUSH}{v,EXCESS(v)}
		\State \textcolor{blue}{P(v):}
		\State $\alpha$ = min\{EXCESS(v), T-OUT(v)[root]\} 
		\State EXCESS(v) = EXCESS(v) - $\alpha$
		\State FIND($\alpha$, k, $\rho$) in T-OUT(v)
		\State \textcolor{blue}{P($e_j = v \rightarrow w$):}
		\If{k' $\leq$ j $\leq$ k' + k}
 		\State UPDATE(j,1) in T-ACCESS(w)
		\State SUM(j,$S_j$) in T-ACCESS(w)
		\If{$j == k$}
			$q_j$ = $\rho$
		\Else
			\State $q_j$ = T-OUT(v)[1,j]
		\EndIf
		\State $f(e_j)$ = $f(e_j)$ + $q_j$
		\State TOTAL(w) = T-IN(w)[root]
		\State UPDATE(head(w) + $S_j$, $q_j$) in T-IN(w)
		\State UPDATE(j, T-OUT(v)[1,j] - $q_j$) in T-OUT(v)
		\State head(w) = head(w) + T-ACCESS(w)[root] 
		\State CLEAR(j) in T-ACCESS(w)
		\State EXCESS(w) = T-IN(w)[root] - TOTAL(w)
		\EndIf
		\State \textcolor{blue}{P(v):} k' = k
		\State \textcolor{blue}{P($e_j = v \rightarrow w$):}
		\If{EXCESS(v) $>$ 0}
			\State block v
		\EndIf
 	\EndFunction
\end{algorithmic}
\end{algorithm}

The \lstinline|RETURN| function given in Algorithm \ref{algo:sv_return} works as follows: In line 3, we find the number $k$ of edges into node $v$ that carry the flow that was pushed to outgoing edges of $v$. In lines 5-9 the excess flow that will be returned on edges $e_j$ s.t. $k<=j<=head(v)$, which brought the excess flow to $v$, is calculated. All those edges $e_j$ will return as much flow as was pushed along them, except for edge $e_k$, which will only return an amount $\rho$. Note that the rest $q_k - \rho$ of the flow pushed along $e_k$ was already forwarded to another node during \lstinline|PUSH|. Lines 10-11 calculate the total amount of flow returned on $e_j$ in this pulse and update the total flow pushed along $e_j$ accordingly. Lines 12-13 calculate the amount of flow returned to node $u$, which is added to the amount of excess flow located at $u$. Finally, in line 17 the $head$ pointer is set to the latest flow quantum that was pushed into $v$ but not returned.

\begin{algorithm}
\caption{Shiloach-Vishkin: RETURN}
\label{algo:sv_return}
\begin{algorithmic}[1]
	\Function{RETURN}{v, EXCESS(v)}
		\State \textcolor{blue}{P(v):}
		\State FIND (T-IN(v)[root] - EXCESS(v), k, $\rho$) in T-IN(v)
		\State EXCESS(v) = 0
		\State \textcolor{blue}{$P(Q_j = (e_j = u \rightarrow v, q_j))$:}
		\State $d_j = q_j$ for $k < j \leq head(v)$
		\State $d_j = q_j - \rho$ if $j == k$
		\State UPDATE(j, 0) in T-IN(v) for $k < j \leq head(v)$
		\State UPDATE(j, $\rho$) in T-IN(v) if $j == k$
		\State UPDATE($p_j$,$d_j$) in T-EDGE($e_j$)
		\State $f(e_j)$ = $f(e_j)$ - T-EDGE($e_j$)[root]
		\State UPDATE($e_j$, T-EDGE($e_j$)[root] in T-SUM(u)
		\State EXCESS(u) = EXCESS(u) + T-SUM(u)[root]
		\State CLEAR($r_j$) in T-EDGE($e_j$)
		\State CLEAR($e_j$) in T-SUM(u)
		\State \textcolor{blue}{P(v):}
		\State head(v) = k
	\EndFunction
\end{algorithmic}
\end{algorithm}

\subsection{Complexity of sequential algorithm}
\label{sec:sv_seq}
The parallel algorithm presented above can easily be turned into a sequential algorithm by executing it with only one processor: This one processor handles all vertices that are unbalanced at the beginning of a pulse, one at a time in an arbitrary order. However, the order of unbalanced vertices of different pulses needs to be enforced. A simple queue thus suffices to implement the set of unbalanced vertices \lstinline|UNBALANCED| used in Algorithm \ref{algo:sv}.

\begin{lemma}
The complexity of the sequential algorithm is $O(n^{3})$\cite{yossi81}.
\end{lemma}
\begin{proof}
All the elementary operations of the algorithm will be associated to either edges or (vertex, pulse) pairs s.t. each edge or pair is assigned only a constant number of elementary operations. As we will show in Section \ref{sec:sv_nr_pulses}, the number of pulses is bounded by $2n$, which implies an $O(n^{2})$\footnote{Note that since $m < n^2$, the $m$ term is dominated by the $n$ term.} bound for the number of (vertex, pulse) pairs for one layered network, yielding a complexity of $O(n^{3})$ for the whole algorithm.

\begin{itemize}
	\item If an edge $e$ becomes saturated during a \emph{PUSH(v, EXCESS(v))}, the resulting operations (which are a constant number) are charged to $e$. In each pulse $p_i$ only one edge $e$ emanating from a vertex $v$ will be used to push along flow without $e$ becoming saturated and thus the elementary operations are charged to the corresponding pair $(v, p_i)$. This implies a bound of $O(n^2)$ elementary operations involved all PUSH operations required to solve the \lstinline|MAX-FLOW| problem in a layered network.
	\item The elementary operations of \emph{RETURN(v, EXCESS(v)} will be charged either to related PUSH operations or again to a (vertex, pulse) pair. When some flow is returned from vertex $v$ at pulse $p_i$, the flow cancels out flow quanta that were pushed to $v$ at some pulse $p_j$ with $j<=i$ and the corresponding push-operation is charged. The returned flow may furthermore cause a change in at most one flow quantum, in which case the elementary operations are assigned to the pair $(v, p_j)$. By the same argument as before, we get a $O(n^{2})$ bound on the number of elementary operations involved in all RETURN operations.
\end{itemize}

These charging rules yield an $O(n^{2})$ bound on the number of elementary operations involved in pushing and returning flow while solving \lstinline|MAX-FLOW| for a layered network, which implies an overall complexity of $O(n^{3}$ for the whole algorithm. 
\end{proof}

\subsection{Processor efficient parallel implementation}
\label{sec:sv_nr_procs}
\begin{lemma}
An allocation of processors to jobs s.t. $n$ processors are sufficient is possible without effecting the complexity of the algorithm. 
\end{lemma}
\begin{proof}
The \textbf{Work-time framework} as presented in \cite{vishkin92} allows for a simple justification of the lemma: the number of rounds is $O(n^2 log n)$: One pulse can be executed in $O(log n)$ time steps (no loops, operations on binary trees only). Since we need at most $2n$ pulses\footnote{see Section \ref{sec:sv_nr_pulses} for the proof.} for solving \lstinline|MAX-FLOW| each of the $O(n)$ layered networks\footnote{according to Dinitz' scheme, as explained in Section \ref{sec:dinitz}.}, the total number of time steps is $T = O(n^{2} log n)$. By similar arguments as we used in Section \ref{sec:sv_seq}, we obtain an amount of total work $W = O(n^{3} log n)$\footnote{see \cite{yossi81} for a more detailed proof}. Applying \emph{Brent's theorem}\cite{Brent74}, we get $\lceil W/p \rceil + T$ rounds, which is $O(n^{2} log n)$ rounds for $p=n$, proving the lemma.
\end{proof}

\subsection{Analysis}
\label{sec:sv_analysis}
We can now provide asymptotic bounds for all steps of Algorithm \ref{algo:sv}: According to Dinitz' scheme, the while-loop is bounded by $O(n)$ iterations. Constructing the residual network $G_r$ can trivially be done in $O(n)$ using n processors, since $m = O(n^{2})$. Breath-first search can be used to construct a layered network $G_l$ out of $G_r$. Efficient parallel implementations of BFS run in $O(m/p +n)$ time using n processors. Adding the found \lstinline|MAX-FLOW| of a layered network to the global \lstinline|MAX-FlOW| can again be done in $O(n)$ time. It remains the calculate the complexity of finding a \lstinline|MAX-FlOW| in a layered network $G_l$, as shown in Algorithm \ref{algo:sv}.

Each \lstinline|PUSH| or \lstinline|RETURN| operation requires $O(log n)$ steps, since they do not contain any loops and statements operate on paths from a leave to the root of a binary tree of height $\lceil log_2(k) \rceil$.

It follows that each round is executed in $O(n*log n)$ time and thus the algorithm overall has a time complexity of $O(n^{2} * log n)$. As shown in Section \ref{sec:sv_pstrees}, a CRCW PRAM is required for the operations on PS-trees. However, Shiloach and Vishkin claim\cite{yossi81}, that a CREW PRAM is sufficient to execute the algorithm within the given bounds, since any concurrent write to a specific memory location will attempt to write the same value. Unfortunately, the paper does not substantiate this claim. 

\section{Algorithm of Goldberg and Tarjan}
\label{sec:goldberg}
The Goldberg-Tarjan algorithm solves \lstinline|MAX-FLOW| on directed acyclic graphs. Contrary to the Shiloach-Vishkin algorithm presented in Section \ref{sec:shiloach}, it is not based on Dinitz' scheme (see Section \ref{sec:dinitz}. The algorithm constructs a blocking flow by moving flow through the network while maintaining a blocking pre-flow. Intuitively, the flow quanta do a depth-first search from $s$ to $t$ in a dynamic graph, where edges become saturated and vertices become blocked.

An \textbf{atom} a is defined as the maximal quantity of flow that so far moved along the same path $P(s,v)$. An atom a can move along forwards on an edge $e \in E$, or backwards along the same edge. Formally, the backward edges are defined as $E^{-1} = \{(w,v)\vert (v,w) \in E \}$. Furthermore, an atom a has the following properties:
\begin{itemize}
	\item \textbf{size(a)}: the amount of excess flow that a carries.
	\item \textbf{position(a)}: the vertex $v \in V$ the atom is currently located at.
	\item \textbf{trace(a)}: The trace, i.e., the sequence of edges the atom a moved so far. $P_t(s, position(a))$ in $E \cup  E^{-1}$. 
	\item \textbf{path(a)}: The simple Path $P_s(s, position(a))$ in $E$. $(v,w) \in P_s$ if $(v,w) \in P_t \and (w,v) \not\in P_t$. In other words, if an atom a traversed an edge $(v,w)$ both forwards and backwards, $(v,w) \in P_t$ but $(v,w) \not\in P_s$.  
\end{itemize}

Algorithm \ref{algo:gt_init} shows the initialization of the Goldberg-Tarjan approach: One atom $a_i$ is created for each edge $e_i=(s, v_i)$ emanating from the start node $s$. Each atom will carry as much flow as can be pushed along the edge  it traverses, thus $size(a_i) = c_(e_i)$. Position and trace of the created atoms are set accordingly.

\begin{algorithm}
\caption{Goldberg-Tarjan: INITIALIZE}
\label{algo:gt_init}
\begin{algorithmic}[1]
	\Function{INITIALIZE}{$N=(V,E,s,t,c)$}
	\ForAll{$e_i = (s,v_i) \in E$}
		\State create new atom $a_i$
		\State $size(a_i) = c(e_i)$
		\State $position(a_i) = v_i$
		\State $trace(a_i) = (s,v_i)$		
	\EndFor
	\EndFunction
\end{algorithmic}
\end{algorithm}

A \emph{phase} of the algorithm is defined as follows:
\begin{itemize}
	\item \textit{phase $1$:} INITIALIZE
	\item \textit{phase $i$:} starts at end of phase $i-1$ and ends as soon as all atoms from phase $i-1$ and all atoms created since the end of phase $i-1$ moved along at least one arc.
\end{itemize}		

After the initialization step, in each subsequent phase $i$ method \lstinline|PUSH| (as depicted in Algorithm \ref{algo:gt_push}) is executed in parallel for all vertices $v \in V \setminus \{s,t\}$ that are not blocked at the beginning of pulse $i$. It moves all atoms $a_i$ currently located at $v$ towards the sink node $s$, if possible, and otherwise towards the start node $s$.

In line 4 of Algorithm \ref{algo:gt_push}, the cumulative sum of flow $S_j$ that atoms $a_i$ located at vertex $v$ carry is calculated. $S(k)$ is the total amount of flow that has to be pushed through outgoing arcs of $v$ that are admissible\footnote{i.e., non-saturated arcs, which can carry some more flow to another nod}. Analogously, the cumulative sum of residual capacities of arcs emanating from $v$ is computed in line 5, with $R(l)$ being the total amount of those residual capacities.

In lines 6-20, each atom $a_i$ is assigned to one or more arc $e_j = (v, w_j)$, s.t.
\begin{itemize}
	\item The sum of flow assigned to an arc $e_j$ is less than or equal to it's residual capacity.
	\item At most one arc is assigned an amount of flow less than the residual capacity of the arc. 
	\item In total, an amount equal to $min\{ECXESS(v), R(l)\}$ is assigned to the set of admissible arcs.
\end{itemize}

If an atom $a_i$ is assigned to more than one arc $(v, w_j)$, it is split, which is done in lines 21-30. If an assignment of an atom $a_i$ to arc $(v, w_j)$ saturates that arc, a new atom $a_k$ is created, carrying the assigned amount of flow along arc $(v, w_j)$. Note that if an assignment of flow to an arc $(v, w_j)$ does not saturate that arc, we do not create a new atom - the original atom $a_i$ will be used to carry the assigned flow. In lines 31-37, all atoms $a_i$ assigned to an arc $(v, w_j)$ will be moved forward along that arc by increasing the flow $f(v, w_j)$ by the amount of flow $size(a_i)$ that atom carries and updating location and trace of $a_i$ accordingly.

After all atoms have been pushed forward, vertex $v$ is blocked if all its outgoing arcs are saturated. Finally, in the last step (lines 42-49) all blocked vertices are checked for remaining atoms $a_i$. Each such atom $a_i$ is then returned on the arc $(v_{a_i}, w)$ that it traversed to enter vertex $v$, which is the last arc on $trace(a_i)$. To do so, the flow $f(v_{a_i}, w)$ is decreased by the amount of flow atom $a_i$ carries and location and trace of $a_i$ are updated.


\begin{algorithm}
\caption{Goldberg-Tarjan: PUSH}
\label{algo:gt_push}
%$\forall v \in V \setminus \{s,t\}$ and $v$ not blocked, do in parallel:
\begin{algorithmic}[1]
	\Function{PUSH}{$v$}
	\State Let $a_1 \dots a_k$ be atoms at vertex $v$
	\State Let $(v,w_1) \dots (v,w_l)$ be admissible arcs at $v$
	\State $\forall 1 \leq j \leq k: S(j) = \Sigma_{i=1}^{j} size(a_i)$
	\State $\forall 1 \leq j \leq l: R(j) = \Sigma_{i=1}^{j} c_r(v,w_i)$
	\State Assign $a_i$ to $(v, w_j)$:
	\If {$S(i) - size(a_i) \geq R(j) - c_r(v, w_j)$}
		\If{$S(i) \leq R(j)$}
			\State assign amount $size(a_i)$ of $a_i$ to $(v,w_j)$
		\Else
			\State assign amount $R(j)-S(i)+size(a_i)$ of $a_i$ to $(v,w_j)$
		\EndIf
	\EndIf
	\If {$S(i) - size(a_i) < R(j) - c_r(v, w_j)$}
		\If{$S(i) > R(j)$}
			\State assign amount $c_r(v,w_j)$ of $a_i$ to $(v,w_j)$
		\Else
			\State assign amount $S(i)-R(j)+c_r(v,w_j)$ of $a_i$ to $(v,w_j)$
		\EndIf
	\EndIf
	\If{$a_i$ assigned to more than 1 $(v,w_j)}$
		\ForAll{$(v,w_j)$ assigned to $a_i$}
			\If{assignment saturates $(v,w_j)$}
				\State create new atom $a_k$ 
				\State $size(a_k$) = quanta assigned to $(v,w_j)$
				\State $trace(a_k) = trace(a_i)$ 
				\State $position(a_k) = position(a_i)$
			\EndIf
		\EndFor
	\EndIf	
	\ForAll{$(v,w_j)$}
		\ForAll{$a_i \leftrightarrow (v,w_j)$}
			\State $f(v,w_j) = f(v,w_j) + size(a_i)$
			\State $location(a_i) = w_j$
			\State $trace(a_i) = trace(a_i) + w_j$
		\EndFor 
	\EndFor
	\If{$\forall (v,w_j):$ saturated}
		\State block $v$
	\EndIf	
	\ForAll{$w \in V$ and $w$ blocked}
		\ForAll{$a_i$ at $w$}
			\State Let $(v_{a_i}, w)$ be last arc on $trace(a_i)$
			\State $f(v_{a_i}, w) = f(v_{a_i}, w) - size(a_i)$
			\State $location(a_i) = v_{a_i}$
			\State $trace(a_i) = trace(a_i) - (v_{a_i}, w)$
		\EndFor
	\EndFor	
	\EndFunction
\end{algorithmic}
\end{algorithm}

\subsection{Analysis}
\label{sec:gt_analysis}
The following observations are important for the analysis of the algorithm:
\begin{itemize}
	\item Atoms can move forward on an edge $e=(v,w) \in E$, increasing the flow $f(v,w)$.
	\item or backward on $e^{-1}=(w,v)$, decreasing $f(v,w)$ and blocking $w$. 
	\item The flow along an edge $f(v,w)$ never becomes negative: an atom can only move backward on $e^{-1}=(w,v)$ if it moved along $e=(v,w)$ first.
	\item $f(v,w)$ increases until $w$ is blocked, after which $f(v,w)$ decreases.
\end{itemize}

\begin{lemma}
The number of created atoms is $\leq m$.
\end{lemma}
\begin{proof}
Atoms created in algorithm
\begin{itemize}
	\item INITIALIZE: $\lvert (s,v_i) \in E \rvert$. $size(a_i) = c(s,v_i)$
	\item PUSH: Splitting an atom: The newly created atom $a'$: $size(a') = c_r(v,w)$ is assigned the residual capacity of an edge $(v,w)$.
\end{itemize}
Clearly, each newly created atom in \lstinline|INITIALIZE| and \lstinline|PUSH| saturates an arc. Each arc may be saturated only once, which implies the claim.
\end{proof}

\begin{lemma}
$\forall a_i:$ at any time, $\lvert trace(a_i) \rvert \leq 2n-3$ arcs.
\end{lemma}
\begin{proof}
An atom $a_i$ only moves backwards on $(v,w) \in E$ once $w$ is blocked. Since $a_i$ moved forwards and backwards on $(v,w)$, $w \notin P_s(a_i)$. $a_i$ never visits $w$ again, since $w$ is now blocked. This implies the following two statements about $trace(a_i$):
\begin{itemize}
	\item $\forall w \in V \setminus \{t\}: trace(a_i)$ contains $(v,w)$ at most once, since an atom $a_i$ may traverse an arc in forward direction at most once. Any arc $(w,t)$ can not be on the trace. Otherwise $a_i$ would have reached the sink node $t$.
	\item $\forall w \in V \setminus \{s,t\}: trace(a_i)$ contains $(w,v)$ at most once, since an atom $a_i$ may traverse an arc in backward direction at most once. Any arc $(t,w)$ can not be on the trace by the same argument as above.  Any arc $(s,w)$ can not be on the trace since this would imply that the an $a_i$ first moved forwards on an arc $(w,s)$.
\end{itemize}
\end{proof}


The definition of phases in Section \ref{sec:goldberg} implies that every atom moves at least one step (forward or backward) during each phase and justifies the following corollary:
\begin{corollary}
The algorithm terminates after at most $2n-3$ phases.
\end{corollary}

By assigning a processor to each atom, each phase requires $O(log n)$ steps and $O(n)$ phases are required. The space complexity of the algorithm is $O(nm)$, which is dominated by storing the traces of $O(m)$ atoms, each of length $O(n$). Thus, the algorithm of Goldberg-Tarjan solves the \lstinline|MAX-FLOW| problem on directed acyclic networks in $O(n*log n)$ time using $m$ processors and $O(nm)$ space on an EREW PRAM\cite{goldberg89}.


\section{Further results and conclusion}
\label{sec:further}
Table \ref{tbl:results} gives an overview of the algorithms presented in this paper as well as some further results. Entries 1-4 depict classical early results, which are the foundation of the Algorithms of Shiloach-Vishkin (5) and Goldberg-Tarjan (6) presented in this paper.  Vishkin (7) combined the two results to an algorithm for an EREW PRAM with time complexity $O(n*logn)$ using $O(n)$ processors and $O(n^{2})$ space. For specialized problems, more efficient algorithms may exist, as shown by Johnson's algorithms solving \lstinline|MAX-FLOW| on planar networks, undirected graphs or for  given maximum flow value in poly-logarithmic time (albeit an unrealistic requirement in number of processors for larger networks). King, Rao and Tarjan (9) showed that the problem can be solved in $O(nm)$ time for larger $m$ on a single processor. A recent result of Orlin (10) proves the conjecture that the \lstinline|MAX-FLOW| problem can be solved in $O(nm)$ time for any $m$ without parallelism. However, this is still slower by a factor of $log n / m$ compared to Vishkin's\cite{vishkin92} 20 year old algorithm for a PRAM, indicating the usefulness of parallel solutions even for problems that are assumed to be inherently sequential, as we showed in Section \ref{sec:cc}.

Although the PRAM model, on which we based our analysis, does not seem to be a "practical" model\footnote{only some prototypes were built so far} of parallel computation, it's high level of abstraction provides the means to develop and analyse parallel algorithms in a clear fashion without the tediousness often involved when working with a more realistic model (such as shared memory or message passing).  Nevertheless, once it was proved that a parallel poly-logarithmic time algorithm does not exist for \lstinline|MAX-FLOW|\footnote{assuming that $P \neq NP$.}, efforts to improve the complexity bounds of PRAM algorithms have somewhat declined.


\begin{table*}
\begin{center}
\begin{tabular}{|l|l|l|l|}
\hline
&Algorithm & Bounds & Remarks \\
\hline
1 & Ford-Fulkerson\cite{ahuja93} & $O((n+m) * f_{max})$ & Sequential algorithm \\
2 & Edmonds-Karp\cite{ahuja93} & $O(n*m^2)$ & Sequential algorithm  \\
3 & Dinitz\cite{dinitz70}& $O(n^2*m)$ & Sequential algorithm \\
  & & & on layered networks\\
4 & Preflow-Push\cite{ahuja93} & $O(n^2*m)$, $O(n*m*log(n^2/m))$ & Various improvements\\
5 & Shiloach-Vishkin\cite{yossi81} & $O(n^2*log m)$, $p=O(n)$ & On CREW PRAM\\
6 & Goldberg-Tarjan\cite{goldberg89} & $O(n*log n)$, $p=O(m)$, $O(nm)$ space & On EREW PRAM \\
7 & Vishkin\cite{vishkin92} & $O(n*log n)$, $p=O(n)$, $O(n^2)$ space & Combines previous two\\
8 & Johnson\cite{johnson87} & $O(log^{3} n)$, $p=O(n^{4})$ or $O(log^{2} n)$, $p=O(n^{6})$ &
For planar networks\\
  & & $	O(log^{2} n)$, $p=O(n^{4})$ & On undirected graphs \\
  & & & or if  $\lvert f_{max} \rvert$ is given\\
9 & King, Rao, Tarjan\cite{King94} & $O(nm)$ & for $m > n^{1+\epsilon}$ \\
10 & Orlin\cite{Orlin13} & $O(nm)$ & for $m \leq O(n^{16/15 - \epsilon})$ \\
\hline
\end{tabular}
\caption{Comparison of different algorithms for \lstinline|MAX-FLOW|}
\label{tbl:results}
\end{center}
\end{table*}

%\pagebreak
\printbibliography
\end{document}
